### 2024-09-07 index.qmd
Fix typo and add RTools and Quarto

---
title: "Code publishing tutorial"
---

## Content

This tutorial covers how to make your code available to others. While the general principles taught in this tutorial apply to most researchers publishing code, we will make very specific suggestions what tools to use.

We will start with creating a simple report which incorporates R code. Then, we will focus on what is required to publish your work:

- removing legal barriers: adding an appropriate license
- high-level documentation: adding a README file
- archival: uploading it to a repository

## Prerequisites

We will assume you have installed [Git](https://git-scm.com/download/), [R](https://cloud.r-project.org/), [RStudio](https://posit.co/download/rstudio-desktop/), and [Quarto](https://quarto.org/docs/download/). Windows users may additionally need [RTools](https://cloud.r-project.org/bin/windows/Rtools/).

If you are really new to R, we recommend looking at the tutorial "[Introduction to R](https://lmu-osc.github.io/introduction-to-R/)". If you have not used Quarto before, we have also prepared a "[Quarto Workshop](https://lmu-osc.github.io/introduction-to-Quarto/)".

[Get started](intro.qmd){.btn-action-primary .btn-action .btn .btn-success .btn-lg}