### 2024-11-27 intro.qmd
Add a background on research compendia to introduction

---
title: "Introduction"
lang: en
---

In the following, we will provide a brief introduction to the concept of _research compendia_.

## The importance of sharing

Suppose you are reading an article about a new imaging method to turn seismological data into subsurface images. The article describes the ideas that went into developing this method and presents a few examples to illustrate its superiority over previous approaches. You got interested and would like to apply this method to your own data. However, with only the article available, it could take months to come up with a working solution, if possible at all. This situation has been put aptly by @Buckheit1995, distilling an idea by Jon Claerbout:

> "An article about computational science in a scientific publication is not the scholarship itself, it is merely advertising of the scholarship. The actual scholarship is the complete software development environment and the complete set of instructions which generated the figures."

Even for articles that are use computers to apply existing methods (rather than reporting on a new method), sharing the source code and being transparent about the computational environment is imperative to making research reproducible [@Ince2012]. By reproducibility, we mean "obtaining consistent results using the same input data; computational steps, methods, and code; and conditions of analysis" [@NASEM2019, p. 46].

## Linking results and computations

This tutorial not only covers sharing the source code, but also connecting it to the results through the creation of dynamic documents. Rather than manually copying numerical results, figures, or tables, they are inserted automatically upon rendering of the article. Dynamic documents bundled together with any necessary data and auxiliary software are called a [research compendium](https://research-compendium.science/) [@Gentleman2007].

The practice to interleave narrative text with code has its roots in the paradigm of literate programming, where documentation and source code are treated as equals and are arranged in a way to maximize understanding [@Knuth1984]. Alternating text and code can be also found in notebook interfaces for exploratory programming, such as provided by Wolfram Mathematica or Jupyter Notebooks [@Kluyver2016], with the added capability of executing the code and embedding its output. With _Sweave_ [@Leisch2002], ideas from both worlds -- literate programming and embedding program output -- were combined into one tool for rendering dynamic documents using the R programming language. It is the predecessor of the R package `knitr` [@Xie2015] which is being used under the hood in this tutorial.^[Specifically, Quarto employs `knitr` to execute chunks of R code.]

Linking results with their computations has benefits for authors and readers. For the author, articles always contain the most recent version of figures, as they are updated automatically when the computation changes. For the readers, it enables understanding exactly how a particular result was obtained if they get access to the underlying research compendium.

## Best practices

When creating the research compendium, there are a few things to consider [by @Arguillas2022, licensed under [CC\ BY\ 4.0](https://creativecommons.org/licenses/by/4.0/)]:

> __Does the research compendium contain everything needed to reproduce a predefined outcome in an organized and parsimonious way?__
> 
> 1. __Completeness:__ The research compendium contains all of the objects needed to reproduce a predefined outcome.
> 2. __Organization:__ It is easy to understand and keep track of the various objects in the research compendium and their relationship over time.
> 3. __Economy:__ Fewer extraneous objects in the compendium mean fewer things that can break and require less maintenance over time.
> 
> __Is descriptive information about the research compendium and its components available and easy to understand?__
> 
> 4. __Transparency:__ The research compendium provides full disclosure of the research process that produced the scientific claim.
> 5. __Documentation:__ Information describing compendium objects is provided in enough detail to enable independent understanding and use of the compendium.
> 
> __Is information about how the research compendium and its components can be used available and easy to understand?__
> 
> 6. __Access:__ It is clear who can use what, how, and under what conditions, with open access preferred.
> 7. __Provenance:__ The origin of the components of the research compendium and how each has changed over time is evident.
> 
> __Is information about the research compendium and its components embedded in code?__
> 
> 8. __Metadata:__ Information about the research compendium and its components is embedded in a standardized, machine-readable code.
> 9. __Automation:__ As much as possible, the computational workflow is script- or workflow-based so that the workflow can be re-executed using minimal actions.
> 
> __Is there a plan for reviewing the research compendium for FAIR and computational reproducibility standards over time?__
> 
> 10. __Review:__ A series of managed activities are needed to ensure continued access to and functionality of the research compendium and its components for as long as necessary.

Although this tutorial guides you through the creation of a research compendium, you are invited to revisit these questions after completion and check whether and how each point was addressed (or not). Further, you can consult them as a checklist for future projects.