### 2024-11-29 about.qmd
Improve license statement on about page

---
title: "About"
author:
  - name: "Florian Kohrt"
    orcid: "0000-0003-0374-5625"
  - name: "Josephine Zerna"
    orcid: "0000-0003-2892-884X"
  - name: "Christoph Scheffel"
    email: "christoph_scheffel@tu-dresden.de"
    orcid: "0000-0001-5963-9229"
bibliography: grateful-refs.bib
---

## Utilized Software

```{r}
#| echo: false

grateful::cite_packages(
  output = "paragraph",
  out.dir = ".",
  pkgs = unique(renv::dependencies(progress = FALSE)$Package),
  passive.voice = TRUE
) |>
  withr::with_options(new = list(warnPartialMatchDollar = FALSE))
```

```{r}
#| output: "asis"
#| echo: false

cat(c("Quarto:", as.character(quarto::quarto_version())))
```

```{r}
#| class.output: "txt code-overflow-scroll"

sessioninfo::session_info()
```

## License and Disclaimer

Except where noted otherwise, the material in this tutorial is licensed under [CC\ BY\ 4.0](https://creativecommons.org/licenses/by/4.0/). None of the discussion in this tutorial constitutes legal advice.

The screenshots of the [RStudio IDE](https://github.com/rstudio/rstudio) are available under the [GNU Affero General Public License v3](https://www.gnu.org/licenses/gpl-3.0.html).

::: {.nte-agplv3 .callout-note collapse="true"}
### GNU Affero General Public License v3

```{.txt .code-overflow-wrap}
Copyright (C) 2024 Posit Software, PBC

This program is free software: you can redistribute it and/or modify it under the terms of version 3 of the GNU Affero General Public License as published by the Free Software Foundation.

This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY; without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the GNU Affero General Public License for more details.

You should have received a copy of the GNU Affero General Public License along with this program.  If not, see <https://www.gnu.org/licenses/>.
```
:::

## Funding

::: {.column-margin}
![](images/bitss.png){width=250px}
:::

This work was partly funded by the Berkeley Initiative for Transparency in the Social Sciences (BITSS), managed by the Center for Effective Global Action (CEGA).