### 2025-01-26 about.qmd
Also mention bibliography file

---
title: "About"
author:
  - name: "Florian Kohrt"
    orcid: "0000-0003-0374-5625"
bibliography: grateful-refs.bib
---

## Utilized Software

```{r}
#| echo: false

grateful::cite_packages(
  output = "paragraph",
  out.dir = ".",
  pkgs = unique(renv::dependencies(progress = FALSE)$Package),
  passive.voice = TRUE
) |>
  withr::with_options(new = list(warnPartialMatchDollar = FALSE))
```

```{r}
#| output: "asis"
#| echo: false

cat(paste0("Quarto ", as.character(quarto::quarto_version()), " was utilized to render the files."))
```

```{r}
#| class-output: "txt code-overflow-scroll"

sessioninfo::session_info()
```

## License and Disclaimer

None of the discussion in this tutorial constitutes legal advice.

Except where noted otherwise, the narrative text in this tutorial is licensed under [CC\ BY-SA\ 4.0](LICENSE-CC-BY-SA.txt); the code without any narrative text is also (at your option) available under [CC0\ 1.0](LICENSE-CC0.txt).

The files `Manuscript.qmd` and `Bibliography.bib` are made available by Josephine Zerna and Christoph Scheffel under [CC0\ 1.0](LICENSE-CC0.txt).

The screenshots of the [RStudio IDE](https://github.com/rstudio/rstudio) are Copyright (C) 2024 Posit PBC and are available under the [GNU Affero General Public License v3](LICENSE-AGPL.txt). The source code is available on [GitHub](https://github.com/rstudio/rstudio/tree/v2024.04.2%2B764).

## Funding

::: {.column-margin}
![](images/bitss.png){width=250px}
:::

This work was partly funded by the Berkeley Initiative for Transparency in the Social Sciences (BITSS), managed by the Center for Effective Global Action (CEGA).