### 2025-02-09 make_readme.qmd

---
title: "Make a README"
engine: knitr
---

## What Does a README Contain?

Having settled on a license, it is time to add a final touch.
Imagine returning to your project in five years,
having forgotten most of the details of what you did exactly.
What would be useful to know in order to quickly understand
what is going on in the project?
This is what needs to be described in the README.
While you could just start writing along,
most READMEs have some sections in common which we describe below.

Name

:   How is the project called?

Badges/Project Status

:   Badges typically report quick facts, like
    [whether the project is actively maintained][repostatus],
    [how many dependencies it has][tinyverse],
    or [whether it has been published in a package repository][metacran].
    Sometimes, the license, any associated DOIs,
    or quality metrics like code coverage by unit tests
    are communicated via badges as well.

[repostatus]: https://www.repostatus.org/
[tinyverse]: https://tinyverse.netlify.app/
[metacran]: https://www.r-pkg.org/services#badges

Description

:   What is the project about? What are its features? Why was it created?

Visuals

:   Is there anything you can show that demonstrates how the project can be used?
    Screenshots or other visuals can make the README more appealing.

Installation/Dependencies

:   What steps need to be taken to run the project? What software needs to be installed?
    R itself and the R packages are already documented as this project uses `renv`.
    Therefore you can focus on all other dependencies,
    such as the system dependencies of R packages
    as well as the version of Quarto.[^renv-quarto]
    Also, don't forget to mention software that you have used for any manual steps.
    See @sec-dependencies for additional information.
    It may make sense to provide these information in the README
    even if you already cite all programs and their version numbers in your manuscript.

[^renv-quarto]: As of August 2024, a proposal for `renv` to record the version of Quarto
has not been implemented, see [rstudio/renv#1143](https://github.com/rstudio/renv/issues/1143).

Usage

:   Which files does the project folder contain? How are they organized?
    How can one run the project -- is there a master script
    or a particular order in which any scripts need to be executed?
    How long does it take to run all scripts?
    Is there additional documentation available?

Support

:   Do you offer support or help, for example, via GitHub discussions
    or a mailing list?

Contributing

:   If the project is active: Can other people contribute? How?
    Do you accept contributions? Do you review issues?
    This section is sometimes outsourced into a file called `CONTRIBUTING.md`.

Authors

:   Who was involved in creating this project? This involves you,
    your co-authors, and anybody you accepted contributions from.

Citation

:   Is there a recommended way to cite this project?
    Is there a published article associated with it
    that you would like to have cited?

License

:   Under which licenses are the works in this project folder available?

## Installation/Dependencies {#sec-dependencies}

An overview over the system dependencies of R packages can be created
using the function `pak::pkg_sysreqs()`.
In combination with `renv`, we can obtain the system dependencies
of all R packages the current project directly depends on:

```{.r filename="Console"}
# First, install pak
renv::install("pak")

# Find all dependencies
deps <- renv::dependencies()$Package |>
   unique() |>
   pak::pkg_deps(dependencies = NA) |>
   getElement("package")

# Identify the system dependencies
pak::pkg_sysreqs(deps)
```

The output may look like the following:

```txt
── Install scripts ────────────────── Fedora 40 ── 
dnf install -y make pandoc git

── Packages and their system dependencies ────────
fs        – make
knitr     – pandoc
remotes   – git
rmarkdown – pandoc
sass      – make
```

We can see that the programs `make`, `pandoc`,
and `git` were identified as system dependencies.
Often, one can obtain their version by running them with the `--version` argument:

```{.bash filename="Terminal"}
make --version
pandoc --version
git --version
```

However, this does not work for all system dependencies.
Specifically, it does not work for libraries -- software that is not supposed to be run on its own.
Identifying their version is beyond the scope of this tutorial.

We also know that we need Quarto to create the PDF,
so let's find out its version as well:

```{.bash filename="Terminal"}
quarto --version
```

If you installed `apaquarto` or any other Quarto extension,
one can query their versions as follows:[^already-included]

[^already-included]: Luckily, the extensions are included in the project folder,
so technically their version is already recorded in the project's files.

```{.bash filename="Terminal"}
quarto list extensions
```

Finally, we know that we installed a $\TeX$ distribution to create the PDF,
so let's find out its version by running:

```{.bash filename="Terminal"}
quarto check
```

The output is quite long and it might look slightly different for you,
but the relevant sections are the following:

```txt
[✓] Checking tools....................OK
      TinyTeX: v2024.09
      Chromium: (not installed)

[✓] Checking LaTeX....................OK
      Using: TinyTex
      Path: /home/r155953/.TinyTeX/bin/x86_64-linux
      Version: 2024
```

Of course, all the system dependencies identified until now
may have dependencies on their own. Use your own judgement to decide when not to dig deeper.

## Create It!

Create your README now as the file `README.md`. If you feel stuck,
you can have a look at the following examples:

::: {#tip-name .callout-tip collapse="true"}
### Name

```{.md .code-overflow-wrap filename="README.md"}
# Penguin Paper

```
:::

::: {#tip-project-status .callout-tip collapse="true"}
### Badges/Project Status

```{.md .code-overflow-wrap filename="README.md"}
[![Project Status: Unsupported – The project has reached a stable, usable state but the author(s) have ceased all work on it. A new maintainer may be desired.](https://www.repostatus.org/badges/latest/unsupported.svg)](https://www.repostatus.org/#unsupported)
```
:::

::: {#tip-description .callout-tip collapse="true"}
### Description

```{.md .code-overflow-wrap filename="README.md"}
This project contains the Quarto manuscript of our study on penguins. It is written in R and uses `renv` to track its dependencies.
```
:::

::: {#tip-installation .callout-tip collapse="true"}
### Installation/Dependencies

``````{.md .code-overflow-wrap filename="README.md"}
## Dependencies

This manuscript requires the following system software to be installed. In addition, we provide the version numbers this manuscript has last been run with:

- [Quarto](https://quarto.org/docs/download/) 1.6.9
- GNU Make 4.4.1
- Pandoc 3.3
- TinyTeX 2024.09
- [R](https://cloud.r-project.org/) 4.4.1

On Fedora Linux, Make and Pandoc can be installed as follows:

```bash
dnf install -y make pandoc
```

Quarto and R can be installed using the links provided. TinyTeX can be installed using Quarto by entering the following into the terminal:

```bash
quarto install tinytex
```

All R packages that this project requires are managed using [`renv`](https://cran.r-project.org/package=renv). Therefore, `renv` needs to be installed first, by entering the following in the R console:

```r
install.packages("renv")
```

Next, one can open a new R session in the root folder of this project and run the following, which should install all required R packages at their recorded versions:

```r
renv::restore()
```
``````
:::

::: {#tip-usage .callout-tip collapse="true"}
### Usage

`````{.md .code-overflow-wrap filename="README.md"}
## Usage

The most important file in this project folder is `Manuscript.qmd` which contains the text of the article as well as the code for its computations. It is accompanied by the following files:

- `Bibliography.bib`: bibliographic references used in the manuscript
- `data.csv`: a data set containing the simplified `palmerpenguins` data
- `data_dictionary.html`: a dictionary to the data file,
  created using `data_dictionary.qmd`

The folder `_extensions` contains the `apaquarto` extension which is used to typeset the PDF accoording to APA guidelines.

The manuscript can be rendered to PDF using the following command:

```bash
quarto render Manuscript.qmd
```
``````
:::

::: {#tip-citation .callout-tip collapse="true"}
### Citation

You need to add your name in the following example:

```{.md .code-overflow-wrap filename="README.md"}
## Citation

Please cite this draft as follows:

> Zerna, Scheffel, & <YOUR NAME> (2024): "A Study on Penguins: A Minimal Reproducible Example". Unpublished manuscript.
```
:::

::: {#tip-citation .callout-tip collapse="true"}
### License

Of course, you would use the same license for the manuscript that you chose in the previous step. You need to add your name in the following example:

```{.md .code-overflow-wrap filename="README.md"}
## License

The manuscript files `Manuscript.qmd`, `Manuscript.tex`, and `Manuscript.pdf` by Josephine Zerna, Christoph Scheffel, and <YOUR NAME> are available under [CC\ BY-SA\ 4.0](https://creativecommons.org/licenses/by-sa/4.0/) or (at your option) under the [AGPLv3](https://www.gnu.org/licenses/agpl-3.0.html) (or later). For further copyright information, see `LICENSE.txt`.
```
:::