### 2025-02-11 about.qmd
Mention all authors and reviewers

---
title: "About"
author:
  - name: "Florian Kohrt"
    orcid: "0000-0003-0374-5625"
  - name: "Josephine Zerna"
    orcid: "0000-0003-2892-884X"
  - name: "Christoph Scheffel"
    orcid: "0000-0001-5963-9229"
bibliography: grateful-refs.bib
---

---

This tutorial was created by Florian Kohrt.
It is based on previous work by Josephine Zerna and Christoph Scheffel.
Comments on earlier versions were kindly provided by
Malika Ihle, Florian Pargent, Jonas Hagenberg, Felix Schönbrodt,
and Sarah von Grebmer zu Wolfsthurn.

## Utilized Software

```{r}
#| echo: false

grateful::cite_packages(
  output = "paragraph",
  out.dir = ".",
  pkgs = unique(renv::dependencies(progress = FALSE)$Package),
  passive.voice = TRUE
) |>
  withr::with_options(new = list(warnPartialMatchDollar = FALSE))
```

```{r}
#| output: "asis"
#| echo: false

cat(paste0("Quarto ", as.character(quarto::quarto_version()), " was utilized to render the files."))
```

```{r}
#| class-output: "txt code-overflow-scroll"

sessioninfo::session_info()
```

## License and Disclaimer

None of the discussion in this tutorial constitutes legal advice.

Except where noted otherwise, the narrative text in this tutorial
is licensed under [CC\ BY-SA\ 4.0](LICENSE-CC-BY-SA.txt);
the code without any narrative text is also (at your option)
available under [CC0\ 1.0](LICENSE-CC0.txt).

The files `Manuscript.qmd` and `Bibliography.bib` are made available
by Josephine Zerna and Christoph Scheffel under [CC0\ 1.0](LICENSE-CC0.txt).

The screenshots of the [RStudio IDE](https://github.com/rstudio/rstudio)
are Copyright (C) 2024 Posit PBC and are available
under the [GNU Affero General Public License v3](LICENSE-AGPL.txt).
The source code is available on [GitHub](https://github.com/rstudio/rstudio/tree/v2024.04.2%2B764).

## Funding

::: {.column-margin}
![](images/bitss.png){width=250px}
:::

This work was partly funded by the Berkeley Initiative for Transparency in the Social Sciences (BITSS), managed by the Center for Effective Global Action (CEGA).