### 2025-02-12 data.qmd
Add helpful code comments

---
title: "Add Data"
engine: knitr
---

## Add Data File

```{r}
#| include: false

invisible(loadNamespace("palmerpenguins")) # Tell renv that we need this package
stopifnot(packageDescription("palmerpenguins", fields = "License") == "CC0")
write.csv(palmerpenguins::penguins, file = "data.csv", row.names = FALSE)
```

You can now download the data set we have prepared for you
and put it into your project folder:

[data.csv](data.csv){.btn .btn-lg .btn-info download=""}

::: {.column-margin}
![palmerpenguins: Palmer Archipelago (Antarctica) Penguin Data](images/palmerpenguins.png){width=250px}
:::

The data set is from the package [`palmerpenguins`][palmerpenguins]
(v`r packageVersion("palmerpenguins")`)
and contains the recorded bill lengths and sex of penguins
living on three islands in the Palmer Archipelago, Antarctica.
It was made available by Allison Horst, Alison Hill, and Kristen Gorman
under the license [CC0\ 1.0][cc0].

[palmerpenguins]: https://allisonhorst.github.io/palmerpenguins/
[cc0]: https://creativecommons.org/publicdomain/zero/1.0/

::: {#imp-legal-restrictions .callout-important}
### Consider Legal Restrictions Before Sharing

Everything you put into the project folder will be shared publicly.
For reasons of reproducibility, this should include the data you analyze.
Of course, you should only share them to the extent you are allowed to.
Besides copyright and similar rights, you need to take into account
applicable privacy laws (e.g., the GDPR for European citizens)
and contractual obligations (e.g., with your data provider).

Privacy laws and contractual obligations may require you
to create a completely anonymized or synthetic data set[^anon-synth-tools] (if possible),
or prohibit any sharing of data, in which case you should provide a reference
to a data repository where they can be obtained from.
For further information, you can watch the talk "[Data anonymity](https://osf.io/j6fy8)"
by Felix Schönbrodt recorded during the LMU Open Science Center Summer School 2023
and have a look at [the accompanying slides](https://osf.io/z6gcu).
Another resource to look at is the presentation
"Creating reproducible packages when data are confidential" by @Vilhuber2024.
:::

[^anon-synth-tools]: For example, using [Amnesia](https://amnesia.openaire.eu/),
[ARX](https://arx.deidentifier.org/), [sdcTools](https://github.com/sdcTools),
or [Synthpop](https://www.synthpop.org.uk/).

## Add Data Dictionary

Whether or not distributing the data set,
it is important to document the meaning (e.g., units)
and values of its variables.
This is typically done with a _data dictionary_ (also called a _codebook_).
In the following, we will demonstrate how to create a simple data dictionary
using the R package [`datawizard`][datawizard].
You can install them now using:

[datawizard]: https://easystats.github.io/datawizard/

```{.r filename="Console"}
renv::install("datawizard")
```

You can put the code that follows in a separate document.
Create it by clicking on _File_ > _New File_  > _Quarto Document..._.
Choose a title such as `Data Dictionary`,
select _HTML_ as format,
uncheck the use of the visual markdown editor, and click on _Create_.
Remove everything except the YAML header (between the `---`).
To make the HTML file self-contained,
also set `embed-resources: true` such that the YAML header looks as follows:

```{.yml filename="data_dictionary.qmd"}
---
title: "Data Dictionary"
format:
  html:
    embed-resources: true
---
```

Then, save it as `data_dictionary.qmd` by clicking on _File_ > _Save_.

To create the actual data dictionary, first write a description for all columns
so others can understand what the variable names mean.
Where necessary, also document their value
-- this is especially important if their meaning is non-obvious.
In the following, we demonstrate this by storing the penguins' binomial name
along with the English name.

``````{cat}
#| engine.opts = list(file = "_data_dictionary.qmd"),
#| class.source = "md",
#| filename = "data_dictionary.qmd"

```{r}
#| echo: false

# Store the description of variables
vars <- c(
  species = "a character string denoting penguin species",
  island = "a character string denoting island in Palmer Archipelago, Antarctica",
  bill_length_mm = "a number denoting bill length (millimeters)",
  bill_depth_mm = "a number denoting bill depth (millimeters)",
  flipper_length_mm = "an integer denoting flipper length (millimeters)",
  body_mass_g = "an integer denoting body mass (grams)",
  sex = "a character string denoting penguin sex",
  year = "an integer denoting the study year"
)

# Store the description of variable values
vals <- list(
  species = c(
    Adelie = "Pygoscelis adeliae",
    Gentoo = "Pygoscelis papua",
    Chinstrap = "Pygoscelis antarcticus"
  )
)
```
``````

Then, load the data and use `datawizard`
to add the descriptions to the `data.frame`:[^not-permanent]

[^not-permanent]: Note that the code provided does not alter the data file
-- no description will be added to `data.csv`.
The descriptions are only added to a (temporary) copy of the data set within R
to create the data dictionary.

::: {.column-margin}
![datawizard: Easy Data Wrangling and Statistical Transformations](images/datawizard.png){width=250px}
:::

``````{cat}
#| engine.opts = list(file = "_data_dictionary.qmd", append = TRUE),
#| class.source = "md",
#| filename = "data_dictionary.qmd"

```{r}
#| echo: false

dat <- read.csv("data.csv")

for (x in names(vars)) {
  if (x %in% names(vals)) {
    dat <- datawizard::assign_labels(
      dat,
      select = I(x),
      variable = vars[[x]],
      values = vals[[x]]
    )
  } else {
    dat <- datawizard::assign_labels(
      dat,
      select = I(x),
      variable = vars[[x]]
    )
  }
}
```

``````

Then, you can create the data dictionary containing the descriptions,
but also some other information about each variable
(e.g., the number of missing values) and print it.

``````{cat}
#| engine.opts = list(file = "_data_dictionary.qmd", append = TRUE),
#| class.source = "md",
#| filename = "data_dictionary.qmd"

```{r}
#| echo: false
#| column: "body-outset"
#| classes: plain

datawizard::data_codebook(dat) |>
  datawizard::data_select(exclude = ID) |>
  datawizard::data_filter(N != "") |>
  datawizard::print_md()
```

``````

```{r}
#| child: "_data_dictionary.qmd"

```

Depending on the type of data, it may also be necessary
to describe sampling procedures (e.g., selection criteria),
measurement instruments (e.g., questionnaires),
appropriate weighting,
already applied preprocessing steps, or contact information.
In our case, as the data has already been published,
we only store a reference to its source.

The data set is from the R package `palmerpenguins`.
If you had it installed
you could use the function `citation()` to create such a reference:

```{r}
#| label: "data-citation"
#| eval: false

citation("palmerpenguins", auto = TRUE) |>
  format(bibtex = FALSE, style = "text")
```

Without the package `palmerpenguins` installed,
you can find a [suggested citation on its website][palmerpenguins-citation]
and add that to your data dictionary:

[palmerpenguins-citation]: https://allisonhorst.github.io/palmerpenguins/#citation

```{r}
#| ref.label = "data-citation",
#| render = function(x, options) gsub("\\n", " ", x = x),
#| echo = FALSE,
#| class.output = "md code-overflow-wrap",
#| attr.output = 'filename="data_dictionary.qmd"'

# This chunk takes the output from the chunk "data-citation"
# and renders it with all newlines replaced by whitespaces.
```

Finally, you can render the data dictionary by running the following:

```{.bash filename="Terminal"}
quarto render data_dictionary.qmd
```

This should create the file `data_dictionary.html`
which you open and view in your web browser.

One could go even further by making the information machine-readable in a standardized way.
We provide an optional example of that in @nte-frictionless.
If you want to learn more about the sharing of research data,
have a look at the tutorial "[FAIR research data management][fair-tutorial]".

[fair-tutorial]: https://lmu-osc.github.io/FAIR-Data-Management/

::: {#nte-frictionless .callout-note collapse="true"}
### Create Machine-Readable Variable Documentation

This example demonstrates how the title and description of the data set,
the description of the variables and their valid values are stored in a machine-readable way.
We'll reuse the descriptions we already created[^value-labels] and add a few others.

[^value-labels]: Unfortunately, the descriptions of values are not reused in this example,
as they are [not supported][enum-labels] by the specification we are using.

[enum-labels]: https://specs.frictionlessdata.io/patterns/#table-schema-enum-labels-and-ordering

First, store the title and description of the data set as a whole:

```{.r filename="Console"}
table_info <- c(
  title = "penguins data set",
  description = "Size measurements for adult foraging penguins near Palmer Station, Antarctica"
)
```

As before, also provide a reference to the source.

```{r}
#| echo: false
#| class-output: "r code-overflow-wrap"
#| attr-output: 'filename="Console"'

# We have provided the data set as CSV file to the readers.
# Therefore, we cannot assume or require that readers have
# the R package palmerpenguins installed. Instead, we create
# the citation on our end and hide how we obtained it.

citation("palmerpenguins", auto = TRUE)$url |>
  paste0("dat_source <- \"", ... = _, "\"") |>
  cat()
```

Next, create a list of the categorical variables' valid values:

```{.r filename="Console"}
valid_vals <- list(
  species = c("Adelie", "Gentoo", "Chinstrap"),
  island = c("Torgersen", "Biscoe", "Dream"),
  sex = c("male", "female"),
  year = c(2007, 2008, 2009)
)
```

Finally, store the descriptions of the variables we already created earlier:

```{.r filename="Console"}
# Store the description of variables
vars <- c(
  species = "a character string denoting penguin species",
  island = "a character string denoting island in Palmer Archipelago, Antarctica",
  bill_length_mm = "a number denoting bill length (millimeters)",
  bill_depth_mm = "a number denoting bill depth (millimeters)",
  flipper_length_mm = "an integer denoting flipper length (millimeters)",
  body_mass_g = "an integer denoting body mass (grams)",
  sex = "a character string denoting penguin sex",
  year = "an integer denoting the study year"
)
```

Generally, metadata are either stored embedded into the data or externally,
for example, in a separate file.
We will use the "[frictionless data](https://frictionlessdata.io/)" standard,
where metadata are stored separately.
Another alternative would be [RO-Crate](https://www.researchobject.org/ro-crate/).

Specifically, one can use the R package [`frictionless`][frictionless]
to create a _schema_ which describes the structure of the data.[^frictionless-note]
For the purpose of the following code,
it is just a nested list that we edit to include our own information.
We also explicitly record in the schema
that missing values are stored in the data file as `NA`
and that the data are licensed under CC0\ 1.0.
Finally, the package is used to create a metadata file that contains the schema.

[frictionless]: https://docs.ropensci.org/frictionless/

[^frictionless-note]: In June 2024, [version 2](https://datapackage.org/)
of the frictionless data standard has been released.
As of November 2024, the R package `frictionless` only supports the first version,
though support for v2 is [planned](https://github.com/frictionlessdata/frictionless-r/labels/datapackage%3Av2).

```{.r filename="Console"}
# Install {frictionless} and the required dependency {stringi}
renv::install(c(
  "frictionless",
  "stringi"
))

# Read data and create schema
dat_filename <- "data.csv"
dat <- read.csv(dat_filename)
dat_schema <- frictionless::create_schema(dat)

# Add descriptions to the fields
dat_schema$fields <- lapply(dat_schema$fields, \(x) {
  c(x, description = vars[[x$name]])
})

# Record valid values
dat_schema$fields <- lapply(dat_schema$fields, \(x) {
  if (x[["name"]] %in% names(valid_vals)) {
    modifyList(x, list(constraints = list(enum = valid_vals[[x$name]])))
  } else {
    x
  }
})

# Define missing values
dat_schema$missingValues <- c("", "NA")

# Create package with license info and write it
dat_package <- frictionless::create_package() |>
  frictionless::add_resource(
    resource_name = "penguins",
    data = dat_filename,
    schema = dat_schema,
    title = table_info[["title"]],
    description = table_info[["description"]],
    licenses = list(list(
      name = "CC0-1.0",
      path = "https://creativecommons.org/publicdomain/zero/1.0/",
      title = "CC0 1.0 Universal"
    )),
    sources = list(list(
      title = "CRAN",
      path = dat_source
    ))
  )
frictionless::write_package(dat_package, directory = ".")
```

This creates the metadata file `datapackage.json` in the current directory.
Make sure it is located in the same folder as `data.csv`,
as together they comprise a [data package](https://specs.frictionlessdata.io/data-package/).
:::

Having added the data and its documentation,
one can view and record the utilized packages with `renv`,
thus bringing the project into a consistent state:

```{.r filename="Console"}
renv::status()
renv::snapshot()
```

## Add Data Citation and Attribution

All data relied upon should be cited in the manuscript
to allow for precise identification and access.
From the "eight core principles of data citation" by @Starr2015,
licensed under [CC0\ 1.0][cc0]:

> **Principle 1 – Importance**: "Data should be considered legitimate,
> citable products of research. Data citations should be accorded the same
> importance in the scholarly record as citations of other research objects,
> such as publications."
> 
> **Principle 3 – Evidence**: "In scholarly literature, whenever and wherever a
> claim relies upon data, the corresponding data should be cited."
> 
> **Principle 5 – Access**: "Data citations should facilitate access to the data
> themselves and to such associated metadata, documentation, code, and other
> materials, as are necessary for both humans and machines to make informed use
> of the referenced data."
> 
> **Principle 7 – Specificity and Verifiability**: "Data citations should
> facilitate identification of, access to, and verification of the specific data
> that support a claim. Citations or citation metadata should include
> information about provenance and fixity sufficient to facilitate verifying
> that the specific time slice, version and/or granular portion of data
> retrieved subsequently is the same as was originally cited."

Now, it's your turn to add an appropriate citation for the data set to the manuscript.
Does your citation adhere to the principles above?

::: {#tip-cite-palmerpenguins .callout-tip collapse="true"}
#### Citing the Data Set (Solution)

You can find an appropriate BibTeX entry on [the package website][palmerpenguins-citation]
or with the function `citation()`:^[Note that this function requires
to have the respective package installed.]

```{r}
#| class.output: "bib code-overflow-scroll"
#| attr-output: 'filename="Bibliography.bib"'

citation("palmerpenguins", auto = TRUE) |>
  transform(key = "horst2020") |>
  toBibtex()
```

Copy the BibTeX entry to the file `Bibliography.bib`.
Then, find the line in the manuscript that says "cite data here"
and replace it with a sentence such as the following:

```{.md filename="Manuscript.qmd"}
The analyzed data are by @horst2020.
```

Render the document to check that the citation is displayed properly.

```{.bash filename="Terminal"}
quarto render Manuscript.qmd
```
:::

While citation happens in the manuscript for reasons of academic integrity and reproducibility,
to comply with any licenses you also may need to provide attribution within your project folder.
Even though the data file we use here does not require attribution,
we recommend adding a short paragraph to `LICENSE.txt`:

```{r}
#| echo: false
#| class-output: "txt code-overflow-wrap"
#| attr-output: 'filename="LICENSE.txt"'

cat(paste0(
  "The penguins data stored in \"data.csv\" by Allison Horst, Alison Hill, and Kristen Gorman available from <",
  citation("palmerpenguins", auto = TRUE)$url,
  "> are licensed under CC0 1.0: <https://creativecommons.org/publicdomain/zero/1.0/>"
))
```

As before, if the license required adding the full license text,
you would also need to copy it to the project folder (if not already in there).

Finally, you can go through the commit routine:

```{.bash filename="Terminal"}
git status
git add .
git commit -m "Add data"
```