### 2025-07-14 copyright.qmd

---
title: "Sharing work of others: Copyright"
engine: knitr
format:
  html:
    include-in-header:
      - text: |
          <style>
          .unstyled-list > ul {
              list-style-type: none;
              padding-left: 1rem;
          }
          .unstyled-list summary {
              cursor: pointer;
          }
          </style>
---

By default, everything you put into the project folder will be shared publicly.
This typically includes:

- **works by others** than yourself or your co-authors,
  because you may use existing data, images or software. 
  Make sure to only share those to the extent you are allowed to.
- **work that was created by yourself** or your co-authors.

Copyright and licensing can be complicated. 
In this chapter, we focus on rather simple (but typical) situations.
For an in-depth explanation of copyright and licenses, 
see our supplementary in-depth material "[Introduction to Copyright and Licenses](in_depth_material/introduction_copyright.qmd)".

## Short Introduction to Copyright and Licenses

Whenever authors create a literary or artistic work
(such as a text, image, video, or software), the copyright law^[in Germany: _Urheberrecht_]
in most countries limits other people from copying, modifying, and sharing it
without the author's express permission. 
Copyright licenses enable authors to free up their works for reuse by others.

A license is a legal document that regulates what other people than the author
can and cannot do with a copyright-protected work, and under which conditions.
We strongly recommend to use *existing boilerplate licenses*
(and not to craft your own license text).

::: {.column-margin}
For the purpose of this chapter, by _license_ we mean _copyright license_.
:::

Works under a free/open license "can be freely studied, applied, copied and/or modified,
by anyone, for any purpose" [@Moller2015].
Importantly, this also means that others do not need to ask or notify the author
and that they can use it also for commercial purposes. If you modified the work and put a sufficient amount of creativity into it,
the result is called a _derivative work_.

Even for free licenses, authors can place certain demands on re-users of their material.
For example, they can (optionally) ...

- demand that they are credited as the creators of a work,
- that the original license must be indicated when a work is reused, 
- that modifications are indicated,
- that derivative works are only shared under the same license, or
- that no further restrictions are imposed on the derivative work.

We explain below how to choose a license that brings the desired "settings".

## Can I Share the Work of Others?

To determine whether a particular work can be included in the project folder
and shared publicly, the first step is to search for its license.
For example, this tutorial has an [About](about.qmd) page detailing its licensing,
whereas other websites sometimes state the license in the footer.

Other places to look for license statements include a project's `README` file or a dedicated `LICENSE` or `COPYING` file.
If a project has been stored in a research repository, sometimes the license is stored in the metadata of the project and displayed somewhere on the project's page.

If you cannot find any license, the work might be
in the **public domain** (see details [here](in_depth_material/introduction_copyright.qmd#public-domain)). In this case, you can simply reuse it.

If no license notice is present and the work is *not* in the public domain,
it is by default  _in copyright_ and you are generally not allowed to share it with others.
You can, however, ask the author to put the corresponding work under a free/open license
-- Wikimedia Commons provides an [email template][commons-templates] to obtain such a permission.

[commons-templates]: https://commons.wikimedia.org/wiki/Commons:Email_templates

Once you found a license, you need to determine whether it is a standard (boilerplate) license and
whether it is a free/open. Then you can share the reused material in your repository.

Here are typical licenses that are free/open:

[osi-list]: https://opensource.org/licenses
[fsf-list]: https://www.gnu.org/licenses/license-list.html#SoftwareLicenses

::: panel-tabset
### Software

For a software license to be considered free/open,
it must be approved by either the [Open Source Initiative][osi-list] (OSI)
or the [Free Software Foundation][fsf-list] (FSF).
Some of the most common free/open software licenses are listed below:

:::: unstyled-list
- <details>
  <summary>GNU Affero General Public License v3.0 (`AGPL-3.0`)</summary>
  __You can share under these conditions__:

  - Attribution
  - Add license text
  - State changes
  - Disclose code
  - Strong copyleft

  [Link to full text](https://www.gnu.org/licenses/agpl-3.0.html#license-text)
  </details>
- <details>
  <summary>GNU General Public License v3.0 (`GPL-3.0`)</summary>
  __You can share under these conditions__:

  - Attribution
  - Add license text
  - State changes
  - Disclose code
  - Strong copyleft

  [Link to full text](https://www.gnu.org/licenses/gpl-3.0.html#license-text)
  </details>
- <details>
  <summary>GNU Lesser General Public License v3.0 (`LGPL-3.0`)</summary>
  __You can share under these conditions__:

  - Attribution
  - Add license text
  - State changes
  - Disclose code
  - Weak copyleft

  [Link to full text](https://www.gnu.org/licenses/lgpl-3.0.html#license-text)
  </details>
- <details>
  <summary>Apache License 2.0 (`Apache-2.0`)</summary>
  __You can share under these conditions__:

  - Attribution
  - Add license text
  - State changes

  [Link to full text](https://www.apache.org/licenses/LICENSE-2.0)
  </details>
- <details>
  <summary>Mozilla Public License 2.0 (`MPL-2.0`)</summary>
  __You can share under these conditions__:

  - Attribution
  - Indicate license
  - Disclose code
  - Weak copyleft

  [Link to full text](https://www.mozilla.org/en-US/MPL/2.0/)
  </details>
- <details>
  <summary>MIT License (`MIT`)</summary>
  __You can share under these conditions__:

  - Attribution
  - Add license text

  [Link to full text](https://choosealicense.com/licenses/mit/#license-text)
  </details>
- <details>
  <summary>BSD 3-Clause "New" or "Revised" License (`BSD-3-Clause`)</summary>
  __You can share under these conditions__:

  - Attribution
  - Add license text

  [Link to full text](https://choosealicense.com/licenses/bsd-3-clause/#license-text)
  </details>
- <details>
  <summary>BSD 2-Clause "Simplified" License (`BSD-2-Clause`)</summary>
  __You can share under these conditions__:

  - Attribution
  - Add license text

  [Link to full text](https://choosealicense.com/licenses/bsd-2-clause/#license-text)
  </details>
::::

If you did not find a particular software license in this list
and consult the resources by the OSI and/or FSF linked above,
be cautious because they also approve special-purpose licenses
that are unfit for software in the general case.
For example, the license CC0\ 1.0 expressively does not grant patent rights
and one should only use software under that license if it is most likely not covered by patents.

### Writing, Images, Audio, Video

Many authors of text and media put their work under a license created by Creative Commons.
Not all Creative Commons licenses are free/open, however, as some of them prohibit commercial use
or the distribution of modified works (such as sharing a translated version).
Therefore, we recommend to only use works made available to you under CC0
or under Creative Commons licenses with the modules "Attribution" (`BY`) or "ShareAlike" (`SA`),
but without the modules "NonCommercial" (`NC`) and "NoDerivatives" (`ND`).
In their most recent version this includes the following licenses:

:::: unstyled-list
- <details>
  <summary>CC0 1.0 (`CC0-1.0`)</summary>
  __You can share under these conditions__: (none)

  [Link to full text][cc0-text]
  </details>
- <details>
  <summary>Creative Commons Attribution 4.0 (`CC-BY-4.0`)</summary>
  __You can share under these conditions__:

  - Attribution
  - Indicate license
  - State changes
  - Anti-DRM

  [Link to full text][cc-by-text]
  </details>
- <details>
  <summary>Creative Commons Attribution-ShareAlike 4.0 (`CC-BY-SA-4.0`)</summary>
  __You can share under these conditions__:

  - Attribution
  - Indicate license
  - State changes
  - Anti-DRM
  - Copyleft

  [Link to full text][cc-by-sa-text]
  </details>
::::

[cc0-text]: https://creativecommons.org/publicdomain/zero/1.0/legalcode#legal-code-title
[cc-by-text]: https://creativecommons.org/licenses/by/4.0/legalcode#legal-code-title
[cc-by-sa-text]: https://creativecommons.org/licenses/by-sa/4.0/legalcode#legal-code-title

Of course, if you encounter a Creative Commons license with a previous version
that satisfies the description above, it can be also considered free/open.
If you encounter a license not from Creative Commons and want to determine whether it is free/open,
you can consult the [lists of licenses accepted at Wikimedia Commons][commons-tags].
Note, however, that some of the licenses there might be inappropriate for text and media.
For example, licenses for software or for documentation commonly require
to distribute a work together with the full license text,
which is impractical for usage in print media.

[commons-tags]: https://commons.wikimedia.org/wiki/Commons:Copyright_tags

### Data

Purely factual data such as measurements are usually not copyrightable,
but literary or artistic works that cross the threshold of originality are.
Additionally, in some jurisdictions, collections of data (i.e., databases)
can be subject to _sui generis_ database rights
which prevent extracting substantial parts of a database.

As a consequence, a database may be licensed differently than its content.
For example, with a database of artworks, the artworks may be licensed individually as chosen by the artists,
but the license for the database as a whole might have been chosen by the curator,
including the selection of entries, field names and any _sui generis_ database rights.
Watch out for any signs that the license for the database and its content are not the same.
In the following, we list some popular free/open data licenses:

:::: unstyled-list
- <details>
  <summary>CC0 1.0 (`CC0-1.0`)</summary>
  __You can share under these conditions__: (none)

  [Link to full text][cc0-text]
  </details>
- <details>
  <summary>Creative Commons Attribution 4.0 (`CC-BY-4.0`)</summary>
  __You can share under these conditions__:

  - Attribution
  - Indicate license
  - State changes
  - Anti-DRM

  [Link to full text][cc-by-text]
  </details>
- <details>
  <summary>Creative Commons Attribution-ShareAlike 4.0 (`CC-BY-SA-4.0`)</summary>
  __You can share under these conditions__:

  - Attribution
  - Indicate license
  - State changes
  - Anti-DRM
  - Copyleft

  [Link to full text][cc-by-sa-text]
  </details>
::::

Note that Creative Commons licenses prior to version 4.0 are not suitable for databases.
You might also encounter the following, less popular free/open data licenses:

:::: unstyled-list
- <details>
  <summary>Open Data Commons Public Domain Dedication & License 1.0 (`PDDL-1.0`)</summary>
  __You can share under these conditions__: (none)

  [Link to full text](https://opendatacommons.org/licenses/pddl/1-0/)
  </details>
- <details>
  <summary>Community Data License Agreement Permissive 2.0 (`CDLA-Permissive-2.0`)</summary>
  __You can share under these conditions__:

  - Add license text

  [Link to full text](https://cdla.dev/permissive-2-0/)
  </details>
- <details>
  <summary>Community Data License Agreement Permissive 1.0 (`CDLA-Permissive-1.0`)</summary>
  __You can share under these conditions__:

  - Attribution
  - Indicate license

  [Link to full text](https://cdla.dev/permissive-1-0/)
  </details>
- <details>
  <summary>Community Data License Agreement Sharing 1.0 (`CDLA-Sharing-1.0`)</summary>
  __You can share under these conditions__:

  - Attribution
  - Indicate license
  - State changes
  - Copyleft

  [Link to full text](https://cdla.dev/sharing-1-0/)
  </details>
::::

Finally, you should know about the following two free/open licenses because they only cover the database,
but not its individual contents or the works produced from it.
If any of these licenses is used, the license for the content must be stated separately.

:::: unstyled-list
- <details>
  <summary>Open Data Commons Open Database License v1.0 (`ODbL-1.0`)</summary>
  __You can share under these conditions__:

  - Attribution
  - Indicate license
  - Anti-DRM
  - Copyleft

  [Link to full text](https://opendatacommons.org/licenses/odbl/1-0/)
  </details>
- <details>
  <summary>Open Data Commons Attribution License v1.0 (`ODC-By-1.0`)</summary>
  __You can share under these conditions__:

  - Attribution
  - Indicate license

  [Link to full text](https://opendatacommons.org/licenses/by/1-0/)
  </details>
::::

If you encounter a license not listed here and want to determine whether it is free/open,
you can consult the [lists of licenses accepted at Wikimedia Commons][commons-tags].
Note, however, that some of the licenses there might be inappropriate for data,
because they do not mention _sui generis_ database rights.
:::

### Under Which Conditions?

Finally, if you verified that the work has been made available to you under a free/open license,
you need to understand the conditions under which you are allowed to include the work in your project folder
and share it with others -- regardless of whether you modified it.

In the following, we will explain the most important conditions of the licenses listed above.
For other licenses we recommend reading their text in full and/or consulting with a lawyer.
The conditions of many popular free/open licenses are also [summarized on ChooseALicense.com][cal-appendix],
which you can use to get a general idea of a license.
Specifically for software licenses the _Open Source Automation Development Lab_
provides [useful checklists detailing the respective obligations][osadl].[^osadl-note]

[cal-appendix]: https://choosealicense.com/appendix/
[osadl]: https://www.osadl.org/Checklists/
[^osadl-note]: Accessing the checklists requires a (free) account on their website.

::: {#nte-private-use .callout-note}
#### Private Use

In the following, we only discuss the conditions under which works may be _shared_.
This is because free/open licenses put no conditions on private use,
where the work is not made available to others.
:::

#### Attribution

Licenses may require that you provide attribution to the author(s)[^others]
and to retain any copyright notices,[^example-notice]
although the specifics vary between licenses.
For example, Creative Commons licenses other than CC0 require
that the name of the author, the title of the work, and a link to the work are provided.

[^others]: or others whose names are provided with the original work
[^example-notice]: Such as `Copyright (c) 2025 Jane Doe`

Projects with many authors often explicitly allow for an attribution
to the collective as a whole rather than being mentioned individually.
Software projects may use copyright notices that only contain the project name,
such as `Copyright (c) 2020-2025 The XYZ Project Contributors` [@Winslow2020].

Where to put the attribution depends on the type of the work,
but in general it is best to keep it close to where the work is used
and to provide all required information. 
For images, for example, this can be a small text next to the image.

#### Modifications of the work

If a work has been modified, some licenses may require that this is indicated (e.g., with a note "adapted from original"). 
Other licenses forbid that a work is modified.
For example, the `NoDerivatives` clause of the `CC-BY-ND-4.0` licenses say:

> If you remix, transform, or build upon the material, you may not distribute the modified material.

Hence, in this case you are free to reuse and redistribute the material - as long as you don't change it.

#### Commercial reuse

The NC clause in the Creative Commons BY-NC (Attribution–NonCommercial) license stands for "NonCommercial", meaning that others are allowed to use, share, adapt, and build upon the work -- but not for commercial purposes. This restriction prohibits any use that is primarily intended for or directed toward commercial advantage or monetary compensation. For example, using the work in an advertisement, selling it, or incorporating it into a product for sale would not be permitted without additional permission from the copyright holder. However, non-profit educational use, personal projects, or sharing within a non-commercial community is generally allowed, as long as proper attribution is given to the original creator.

#### Indicate License / Add License Text

Licenses may either require that one mentions the applicable license
by name and possibly link ("Indicate license")
or they even require to copy the full license text along with the used work ("Add license text").

#### Copyleft

If the original work has been made available under a "copyleft" license, 
any derivative work that is shared with others
must be put under the same license as the original work[^or-compatible]
-- or you commit a copyright violation.

This explains the origin of the term _copyleft_,
as it "turns copyright around" to make works permanently free.
In the realm of Creative Commons licenses, copyleft is also called _share alike_ (e.g. in the `CC-BY-SA-4.0` license).

[^or-compatible]: or sometimes also under a later version or under a compatible license

What is considered creation of a derivative work varies between jurisdictions,
but the following are typical examples:
translating a text into a different language,
taking a picture of an artwork,
adding a song to a video, adapting a computer program to fit own use cases,
or taking a [screenshot of a computer program][commons-screenshots].
In all of these cases, if the original work is under a copyleft license,
then sharing of the resulting derivative work is only allowed
if it is put under the same license.

[commons-screenshots]: https://commons.wikimedia.org/wiki/Commons:Screenshots

The question is then: Which uses of a work do not create a derivative work?
In many cases, if an unaltered excerpt "is used to illuminate an idea
or provide an example in another larger work" [@CC2024NoDerivs] no derivative work is made.
The excerpt needs to "remain separate and distinct in the new context" [@Kreutzer2024, p. 44]
and not be "merged with other material into a new and larger work" [@Kreutzer2024, p. 45].
For example, embedding images in a document and clearly separating them from the rest
or assembling works as part of a collection typically would not create a derivative work.
However, due to a lack of case law, there is only little guidance
on what constitutes a derivative work in general.

Also note that the copyleft licenses we discuss here do not mandate sharing.
Copyleft (and attribution) clauses are only triggered *if* the work is shared [@CC2015].
This means that if you only use a work internally, you do not need to share your derivative works.

::: {#tip-cc-licenses-checklist .callout-tip}
#### Required Information under Creative Commons Licenses

For most Creative Commons licenses,
providing the following information is required (if available):

- [ ] Original author name
- [ ] Copyright notice
- [ ] Title of the work
  - [ ] Link to the work
- [ ] License name
  - [ ] Link to the license
- [ ] Link to warranty disclaimer
- [ ] Indicate modification
:::

Obviously, if you are in doubt whether a condition of the license applies in your use case,
it is better to either follow it or to ask the author for clarification or explicit permission.

We recommend that you always provide attribution for all works under a free/open license
-- even if the license does not strictly require it (e.g., CC0).
This also follows from the rules of good scientific practice (and not only from copyright law).

::: {#imp-non-copyright-restrictions .callout-important}
#### Non-copyright Restrictions

The free/open licenses we discuss here mostly grant copyrights
(with some of them explicitly granting patent rights and _sui generis_ database rights).
Therefore, you may lack other rights such as personality, privacy, moral, or trademark rights.
For example, [sharing photos that depict people][commons-photographs]
is not only a matter of copyright, but also of privacy rights.
Also, their commercial use may require the consent of the depicted person.
:::

[commons-photographs]: https://commons.wikimedia.org/wiki/Commons:Photographs_of_identifiable_people

## Practical Exercise: Adding an Image

Let's practice what you learned by adding an image to the manuscript.
We'll use [this picture of a penguin from Flickr](https://flic.kr/p/2pEKnUr).
First, try to answer the following questions:

1. Under which license is the picture available?
2. Is it free/open?
3. Under which conditions may the work be shared?

::: {#tip-example-license .callout-tip collapse="true"}
### License of Penguin Picture (Solution)

1. The picture has been put under the license [CC0\ 1.0][cc0]
2. The license is free/open
3. The license contains no (major) conditions
:::

[cc0]: https://creativecommons.org/publicdomain/zero/1.0/

You can now download the image
and move it to your project folder (maybe call it `penguin.jpg`).
In your manuscript file, find the line that says "add image here", remove it,
and add the image instead, including proper attribution.

::: {#tip-example-license .callout-tip collapse="true"}
### Attribution for Penguin Picture (Solution)

Although the license CC0 doesn't require it,
we still recommend providing proper attribution.

- Author name: `steve b`
- Copyright notice: (not available)
- Title of the work: `Chinstrap Penguin`
  - Link to the work: <https://www.flickr.com/photos/192320315@N02/53603388557/> or <https://flic.kr/p/2pEKnUr>
- License name: `CC0 1.0`
  - Link to the license: <https://creativecommons.org/publicdomain/zero/1.0/>
- Link to warranty disclaimer: (not available)
- Indicate modification: (no existing modification notices)

```{.md .code-overflow-wrap}
!["[Chinstrap Penguin](https://flic.kr/p/2pEKnUr)" by "steve b" licensed under [CC0 1.0](https://creativecommons.org/publicdomain/zero/1.0/)](penguin.jpg){width=50%}
```

If it is likely that the article gets printed,
you might want to ensure that the link to the work is retained in that case:

```{.md .code-overflow-wrap}
!["[Chinstrap Penguin](https://flic.kr/p/2pEKnUr)" by "steve b" available under [CC0 1.0](https://creativecommons.org/publicdomain/zero/1.0/) from [`flic.kr/p/2pEKnUr`](https://flic.kr/p/2pEKnUr)](penguin.jpg){width=50%}
```
:::

Let's render the document and see if the changes were successful
(open `Manuscript.pdf` afterwards):

```{.bash filename="Terminal"}
quarto render Manuscript.qmd
```

In addition, we recommend creating a file called `LICENSE.txt`
to describe the license of the image.
Click on _File_ > _New File_ > _Text File_,
then on _File_ > _Save_, and choose `LICENSE.txt` as filename.

```{.txt .code-overflow-wrap filename="LICENSE.txt"}
"Chinstrap Penguin" stored in "penguin.jpg" by "steve b" available from <https://flic.kr/p/2pEKnUr> is licensed under CC0 1.0: <https://creativecommons.org/publicdomain/zero/1.0/>
```

If the license required adding the full license text,
you would also need to copy it to the project folder.

Finally, you can store the changes in the version control system:

```{.bash filename="Terminal"}
git status
git add .
git commit -m "Add image"
```

## Wrap-up

If you would like to learn more about copyright and licenses
you might find the following resources interesting:

- Our own [in-depth material on copyright and licenses](in_depth_material/introduction_copyright.qmd) (which is an extension of this chapter)
- "Open Content -- A Practical Guide to Using Creative Commons Licences"
  by @Kreutzer2024
- "Creative Commons Certificate for Educators, Academic Librarians,
  and Open Culture" by @CC2024Certificate
- "Freie Software -- Zwischen Privat- und Gemeineigentum" by @Grassmuck2004
- "Rechtsfragen bei Open Science: Ein Leitfaden" by @Kreutzer2021