### 2025-07-14 in_depth_material/introduction_copyright.qmd

---
title: "Introduction to Copyright and Licensing"
engine: knitr
format:
  html:
    include-in-header:
      - text: |
          <style>
          .unstyled-list > ul {
              list-style-type: none;
              padding-left: 1rem;
          }
          .unstyled-list summary {
              cursor: pointer;
          }
          </style>
---

By default, everything you put into the project folder will be shared publicly.
In many instances, this will also include works by others than yourself or your co-authors,
because you may use existing data, images or software.
In these cases, you should make sure to only share them to the extent you are allowed to.
In the following, we will be focusing on copyright,
but of course, there can also be other restrictions.

::: {#nte-copyright-tldr .callout-note}
## Copyright TL;DR

Every expression of an idea, such as a literary or an artistic works,
is automatically covered by copyright.
Put simply, this means that nobody else than the copyright holder
is allowed to copy, modify, or share it.
Copyright licenses can grant others some of the necessary rights,
but mostly don't cover publicity, privacy, moral, patent, or trademark rights.
Certain exceptions and limitations to copyright (such as fair use or the right to quote)
may obviate the need for a license.

With regard to copyright, free/open licenses can give permission to share a work by others.
These licenses often require to attribute the authors,
indicate whether changes were made,
and provide the text of the license, among other things.

The following video by @Brinken2021
embedded from the TIB AV-Portal ([Data Privacy Statement](https://av.tib.eu/privacy))
explains the correct attribution of works under Creative Commons licenses.

```{=html}
<details>
  <summary>Load video</summary>
  <p><div style="position: relative; width: 100%; aspect-ratio: 16 / 9;">
    <iframe loading="lazy" src="https://av.tib.eu/player/53550" allowfullscreen style="position: absolute; top: 0; left: 0; width: 100%; height: 100%;"></iframe>
  </div></p>
  <p>The video “The correct indication of a license: Aspects to consider when reusing the work of others” by Helene Brinken, Jonas Hauss, and Jessika Rücknagel is licensed under <a href="https://creativecommons.org/licenses/by/3.0/de/deed.de">CC&nbsp;BY&nbsp;3.0&nbsp;DE</a>.</p>
</details>
```
:::

## Introduction to Copyright and Licenses

Whenever authors create a literary or artistic work
(such as a text, image, video, or software), the copyright law^[in Germany: _Urheberrecht_]
in most countries limits other people from copying, modifying, and sharing it
without the author's express permission.
This even applies if the author makes it available to others (e.g., on their website):
First and foremost others are not allowed to copy, modify, or share it.
Importantly, this legal default of "all rights reserved" was created to benefit publishers,
not authors [@Fogel2006], and runs counter to many cultural and scientific processes.
Copyright licenses enable authors to free up their works for reuse by others.

Note that copyright law [varies between countries][commons-rules]
and the rule of territoriality (_lex loci protectionis_) applies:
Applicable are the laws of the country where a work is used.
Also, if the work was created as part of the author's job,
it might be their employer who holds the copyright to the work,
depending on the country and contract.

[commons-rules]: https://commons.wikimedia.org/wiki/Commons:Copyright_rules_by_territory

A license is a legal document that regulates what other people than the author
can and cannot do with a copyright-protected work, and under which conditions.
The license may be exclusive, such that it is only granted to one recipient,
effectively limiting the author in dealing with their own work.
It can also be non-exclusive, thus giving rights to others without the author losing them.
To avoid that every author who would like to share their work writes their own license,
pre-formulated "boilerplate" licenses have been created.
For example, you might have heard about ["Creative Commons" licenses][cc-licenses],
which are used by Wikipedia.
Recognizing them facilitates both sharing (on the side of the author)
and re-using (on the side of the recipient).^[In fact, authors should not
create their own license texts if the purpose is to share their work freely
and openly with others.
By using a boilerplate license others know what to expect,
also because it may have already been [tested][sfconservancy] in [courts][cc-db].]

[cc-licenses]: https://creativecommons.org/share-your-work/cclicenses/
[sfconservancy]: https://sfconservancy.org/copyleft-compliance/
[cc-db]: https://legaldb.creativecommons.org/

::: {.column-margin}
For the purpose of this tutorial, by _license_ we mean _copyright license_.
:::

An important class of boilerplate licenses are _free/open_ licenses.
Put simply, works under a free/open license "can be freely studied, applied, copied and/or modified,
by anyone, for any purpose" [@Moller2015].
Importantly, this also means that others do not need to ask or notify the author
and that they can use it for commercial purposes.
Coming back to the example of Creative Commons licenses,
not all of them are free/open.
For example, some restrict commercial use and modification.
The Creative Commons license "[Attribution-ShareAlike 4.0][cc-by-sa]"
is an example of a free/open license, however,
and we will cover several others below.

[cc-by-sa]: https://creativecommons.org/licenses/by-sa/4.0/

By the choice of license, authors can demand that they are credited,
that the original license is indicated, that modifications are indicated,
that derivative works are only shared under the same license,
and that no further restrictions are imposed on the work.
Software licenses may additionally require to make the source code available
to everybody the software is shared with
and often require to display the full text of the license upon distribution.

::: {#nte-terminology .callout-note collapse="true"}
### A Note on Terminology

The terms _free_ and _open_, especially with regard to software, come with a history.
The Free Software Foundation (FSF) was founded in 1985
to protect "four essential freedoms" [-@FSF2024] of a program's user.
These are the freedoms to use, study, share, and improve a program.
Software whose users legally and practically have these freedoms
(because, among other things, they have access to its source code) is considered _free_.
The four freedoms are seen as vital for a society as a whole
in the sense that they enable sharing, cooperation and ultimately freedom in general.
The FSF maintains [a list of software licenses][fsf-list]
that it considers to be protecting the four freedoms.
Sometimes the term _libre_ (Spanish and French for _free_)
is used to make a distinction from _gratis_ software.
You can learn more about free software at [Write Free Software](https://writefreesoftware.org/learn).

[fsf-list]: https://www.gnu.org/licenses/license-list.html#SoftwareLicenses

The Open Source Initiative (OSI), which was founded in 1998, follows a more pragmatic approach.
It is concerned with developing high-quality software,
for which everyone's ability to obtain, modify and contribute back the source code
is considered beneficial.
Access to the source code is one out of multiple conditions for software
to be considered _open source_ by the OSI [-@OSI2007],
which equally maintains [a list of approved licenses][osi-list].

[osi-list]: https://opensource.org/licenses

Access to the source code is a necessary, but not sufficient requirement
both for _free software_ and for _open source software_.
Conversely, both can (and frequently are) sold for money,
as their respective criteria only apply once one has access to the software.
Throughout this tutorial, we write "free/open license"
to mean a license that is approved by either the FSF or the OSI.
Software which is neither but makes available its source code
is sometimes referred to as _source-available software_.

Apart from software, the term _open access_ has often been used for works
that are available at no cost.
For example, this is the commonality of bronze, green, hybrid, gold,
and diamond/platinum open access _articles_,
which otherwise vary in the rights that are granted to readers.
In 2002, the Budapest Open Access Initiative declared
that _open access_ additionally includes the right to use articles for any purpose,
and in 2003, the Bethesda Statement and the Berlin Declaration added
the right to make derivative works.

Two other notable definitions include the _Open Definition_ [@OKFN2016],
which was first drafted in 2005, and the definition of _Free Cultural Works_ [@Moller2015],
for which the open editing phase began in 2006.
They are largely viewed as compatible with one another.

:::: {.column-margin}
![](images/Definition_of_Free_Cultural_Works.svg){width=250px}
::::
:::

### Can I Share That Work?

To determine whether a particular work can be included in the project folder
and shared publicly, the first step is to search for its license.
For example, this tutorial has an [about page](../about.qmd) detailing its licensing,
whereas other websites sometimes state the license in the footer.
Other places to look include a project's `README` file or a dedicated `LICENSE` or `COPYING` file.
If a project has been stored in a research repository,
yet another means of communicating its copyright status might exist.
For media from Wikimedia Commons, an [Attribution Generator][commons-gen] exists.

[commons-gen]: https://lizenzhinweisgenerator.de/?lang=en

::: {#tip-private-use .callout-tip}
#### Consider Complicated Licensing Situations

For works by one or only few authors, it is possible
that one license consistently applies to the whole content.
More often than not, however, the situation is a little more complicated:

- Different parts of a work may be available under different licenses:
  A document may incorporate direct quotes or images by others.
  In this case, the license of the document may not cover those.
  In the best case, the authors indicate which parts of the work are covered by which license.
- A work may be available under different licenses at the choice of the (re-)user.
  This is also called multi-licensing.
  For example, authors might allow to use a work under any version of a license.
- In contrast, it may also be the case that one has to comply
  with the conditions of multiple licenses at the same time.
  This is the case, for example, if a derivative work is published
  under a different license than the original work.
:::

#### Public domain

If you cannot find any license, the work might be
in the public domain^[in Germany: _Gemeinfreiheit_] depending on national law.
For example, the work may not cross the [threshold of originality][commons-threshold],
it may not [qualify for copyright protection][commons-unprotected]
-- such as official works by the government --,
or the copyright protection might have expired.
According to the [Berne convention][berne-convention],
copyright expires at earliest 50 years after the author's death,
but in [many countries][commons-pd] it expires even later
-- you can use a [public domain calculator][pd-calc] to check.
If no license notice is present and the work is not in the public domain,
it is _in copyright_ and you are generally not allowed to share it with others.
You can, however, ask the author to put the corresponding work under a free/open license
-- Wikimedia Commons provides an [email template][commons-templates] to obtain such a permission.
Without permission you can check your national laws
for limitations and exceptions to copyright.^[in Germany: _Schranken des Urheberrechts_]
Examples include the right to quote, [trivial use][commons-minimis],
and the US-American [fair use][wiki-fair-use].

[commons-threshold]: https://commons.wikimedia.org/wiki/Commons:Threshold_of_originality
[commons-unprotected]: https://commons.wikimedia.org/wiki/Commons:Unprotected_works
[berne-convention]: https://en.wikipedia.org/wiki/Berne_Convention
[commons-pd]: https://commons.wikimedia.org/wiki/Help:Public_domain#Copyright_terms_by_country
[pd-calc]: https://lawflow.org/public-domain-calculators/
[commons-templates]: https://commons.wikimedia.org/wiki/Commons:Email_templates
[commons-minimis]: https://commons.wikimedia.org/wiki/Commons:De_minimis
[wiki-fair-use]: https://en.wikipedia.org/wiki/Wikipedia:FAQ/Copyright#What_is_fair_use?

Having found a license, you need to determine
(1) whether it is a standard (boilerplate) license,
(2) whether it is a free/open,
and (3) whether it is appropriate for the particular type of work and use case.

1. __Standard licenses__ are usually stated by their name (or an abbreviation),
   which may include a version number.
   Most standard licenses were assigned
   an [SPDX identifier](https://spdx.dev/ids) for unambiguous identification,
   though the identifier is often not indicated for a particular work
   (which would be best practice).
   Also, not all standard licenses have SPDX identifiers.
   You can view a [list of all licenses with SPDX identifiers here](https://spdx.org/licenses/).
2. __Free/open__: We have compiled lists of the most common free/open licenses below,
   categorized by the type of work they apply to.
   Be sure to look for an exact match,
   as there are often many licenses with very similar names.
   If a particular license is not listed below, we also provide links to other resources
   which you can use to determine whether a license is free/open.
3. __Type of work and use case__:
   Many licenses were written with a particular type of work in mind,
   along with a certain use case.
   Software licenses, for example, may consider that the source code
   is the preferred form for making modifications
   and that software can be encumbered by patents,
   while licenses for data can differentiate between the database
   and any works produced from it.
   Take caution if a license applies to a type work for which it was not originally created
   and conflicts with your use case.
   Sometimes, the license name hints at the intended type of work,
   but when in doubt always read the full license text (and/or consult with a lawyer)
   to determine whether the conditions are acceptable to you.

::: panel-tabset
### Software

For a software license to be considered free/open,
it must be approved by either the [Open Source Initiative][osi-list] (OSI)
or the [Free Software Foundation][fsf-list] (FSF).
Some of the most common free/open software licenses are listed below:

:::: unstyled-list
- <details>
  <summary>GNU Affero General Public License v3.0 (`AGPL-3.0`)</summary>
  __Major conditions when sharing works__:

  - Attribution
  - Add license text
  - State changes
  - Disclose code
  - Strong copyleft

  [Link to full text](https://www.gnu.org/licenses/agpl-3.0.html#license-text)
  </details>
- <details>
  <summary>GNU General Public License v3.0 (`GPL-3.0`)</summary>
  __Major conditions when sharing works__:

  - Attribution
  - Add license text
  - State changes
  - Disclose code
  - Strong copyleft

  [Link to full text](https://www.gnu.org/licenses/gpl-3.0.html#license-text)
  </details>
- <details>
  <summary>GNU Lesser General Public License v3.0 (`LGPL-3.0`)</summary>
  __Major conditions when sharing works__:

  - Attribution
  - Add license text
  - State changes
  - Disclose code
  - Weak copyleft

  [Link to full text](https://www.gnu.org/licenses/lgpl-3.0.html#license-text)
  </details>
- <details>
  <summary>Apache License 2.0 (`Apache-2.0`)</summary>
  __Major conditions when sharing works__:

  - Attribution
  - Add license text
  - State changes

  [Link to full text](https://www.apache.org/licenses/LICENSE-2.0)
  </details>
- <details>
  <summary>Mozilla Public License 2.0 (`MPL-2.0`)</summary>
  __Major conditions when sharing works__:

  - Attribution
  - Indicate license
  - Disclose code
  - Weak copyleft

  [Link to full text](https://www.mozilla.org/en-US/MPL/2.0/)
  </details>
- <details>
  <summary>MIT License (`MIT`)</summary>
  __Major conditions when sharing works__:

  - Attribution
  - Add license text

  [Link to full text](https://choosealicense.com/licenses/mit/#license-text)
  </details>
- <details>
  <summary>BSD 3-Clause "New" or "Revised" License (`BSD-3-Clause`)</summary>
  __Major conditions when sharing works__:

  - Attribution
  - Add license text

  [Link to full text](https://choosealicense.com/licenses/bsd-3-clause/#license-text)
  </details>
- <details>
  <summary>BSD 2-Clause "Simplified" License (`BSD-2-Clause`)</summary>
  __Major conditions when sharing works__:

  - Attribution
  - Add license text

  [Link to full text](https://choosealicense.com/licenses/bsd-2-clause/#license-text)
  </details>
::::

If you did not find a particular software license in this list
and consult the resources by the OSI and/or FSF linked above,
be cautious because they also approve special-purpose licenses
that are unfit for software in the general case.
For example, the license CC0\ 1.0 expressively does not grant patent rights
and one should only use software under that license if it is most likely not covered by patents.

### Writing, Images, Audio, Video

Many authors of text and media put their work under a license created by Creative Commons.
Not all Creative Commons licenses are free/open, however, as some of them prohibit commercial use
or the distribution of modified works (such as sharing a translated version).
Therefore, we recommend to only use works made available to you under CC0
or under Creative Commons licenses with the modules "Attribution" (`BY`) or "ShareAlike" (`SA`),
but without the modules "NonCommercial" (`NC`) and "NoDerivatives" (`ND`).
In their most recent version this includes the following licenses:

:::: unstyled-list
- <details>
  <summary>CC0 1.0 (`CC0-1.0`)</summary>
  __Major conditions when sharing works__: (none)

  [Link to full text][cc0-text]
  </details>
- <details>
  <summary>Creative Commons Attribution 4.0 (`CC-BY-4.0`)</summary>
  __Major conditions when sharing works__:

  - Attribution
  - Indicate license
  - State changes
  - Anti-DRM

  [Link to full text][cc-by-text]
  </details>
- <details>
  <summary>Creative Commons Attribution-ShareAlike 4.0 (`CC-BY-SA-4.0`)</summary>
  __Major conditions when sharing works__:

  - Attribution
  - Indicate license
  - State changes
  - Anti-DRM
  - Copyleft

  [Link to full text][cc-by-sa-text]
  </details>
::::

[cc0-text]: https://creativecommons.org/publicdomain/zero/1.0/legalcode#legal-code-title
[cc-by-text]: https://creativecommons.org/licenses/by/4.0/legalcode#legal-code-title
[cc-by-sa-text]: https://creativecommons.org/licenses/by-sa/4.0/legalcode#legal-code-title

Of course, if you encounter a Creative Commons license with a previous version
that satisfies the description above, it can be also considered free/open.
If you encounter a license not from Creative Commons and want to determine whether it is free/open,
you can consult the [lists of licenses accepted at Wikimedia Commons][commons-tags].
Note, however, that some of the licenses there might be inappropriate for text and media.
For example, licenses for software or for documentation commonly require
to distribute a work together with the full license text,
which is impractical for usage in print media.

[commons-tags]: https://commons.wikimedia.org/wiki/Commons:Copyright_tags

### Data

Purely factual data such as measurements are usually not copyrightable,
but literary or artistic works that cross the threshold of originality are.
Additionally, in some jurisdictions, collections of data (i.e., databases)
can be subject to _sui generis_ database rights
which prevent extracting substantial parts of a database.

As a consequence, a database may be licensed differently than its content.
For example, with a database of artworks, the artworks may be licensed individually as chosen by the artists,
but the license for the database as a whole might have been chosen by the curator,
including the selection of entries, field names and any _sui generis_ database rights.
Watch out for any signs that the license for the database and its content are not the same.
In the following, we list some popular free/open data licenses:

:::: unstyled-list
- <details>
  <summary>CC0 1.0 (`CC0-1.0`)</summary>
  __Major conditions when sharing works__: (none)

  [Link to full text][cc0-text]
  </details>
- <details>
  <summary>Creative Commons Attribution 4.0 (`CC-BY-4.0`)</summary>
  __Major conditions when sharing works__:

  - Attribution
  - Indicate license
  - State changes
  - Anti-DRM

  [Link to full text][cc-by-text]
  </details>
- <details>
  <summary>Creative Commons Attribution-ShareAlike 4.0 (`CC-BY-SA-4.0`)</summary>
  __Major conditions when sharing works__:

  - Attribution
  - Indicate license
  - State changes
  - Anti-DRM
  - Copyleft

  [Link to full text][cc-by-sa-text]
  </details>
::::

Note that Creative Commons licenses prior to version 4.0 are not suitable for databases.
You might also encounter the following, less popular free/open data licenses:

:::: unstyled-list
- <details>
  <summary>Open Data Commons Public Domain Dedication & License 1.0 (`PDDL-1.0`)</summary>
  __Major conditions when sharing works__: (none)

  [Link to full text](https://opendatacommons.org/licenses/pddl/1-0/)
  </details>
- <details>
  <summary>Community Data License Agreement Permissive 2.0 (`CDLA-Permissive-2.0`)</summary>
  __Major conditions when sharing works__:

  - Add license text

  [Link to full text](https://cdla.dev/permissive-2-0/)
  </details>
- <details>
  <summary>Community Data License Agreement Permissive 1.0 (`CDLA-Permissive-1.0`)</summary>
  __Major conditions when sharing works__:

  - Attribution
  - Indicate license

  [Link to full text](https://cdla.dev/permissive-1-0/)
  </details>
- <details>
  <summary>Community Data License Agreement Sharing 1.0 (`CDLA-Sharing-1.0`)</summary>
  __Major conditions when sharing works__:

  - Attribution
  - Indicate license
  - State changes
  - Copyleft

  [Link to full text](https://cdla.dev/sharing-1-0/)
  </details>
::::

Finally, you should know about the following two free/open licenses because they only cover the database,
but not its individual contents or the works produced from it.
If any of these licenses is used, the license for the content must be stated separately.

:::: unstyled-list
- <details>
  <summary>Open Data Commons Open Database License v1.0 (`ODbL-1.0`)</summary>
  __Major conditions when sharing works__:

  - Attribution
  - Indicate license
  - Anti-DRM
  - Copyleft

  [Link to full text](https://opendatacommons.org/licenses/odbl/1-0/)
  </details>
- <details>
  <summary>Open Data Commons Attribution License v1.0 (`ODC-By-1.0`)</summary>
  __Major conditions when sharing works__:

  - Attribution
  - Indicate license

  [Link to full text](https://opendatacommons.org/licenses/by/1-0/)
  </details>
::::

If you encounter a license not listed here and want to determine whether it is free/open,
you can consult the [lists of licenses accepted at Wikimedia Commons][commons-tags].
Note, however, that some of the licenses there might be inappropriate for data,
because they do not mention _sui generis_ database rights.
:::

::: {#cau-license-versions .callout-caution}
### License Versions Are Important

You may have noticed that in many cases a version number is part of the license name.
This is because the organizations that created the licenses sometimes publish updated versions
to accommodate for developments in copyright law and the communities that use the licenses.
For example, the Creative Commons licenses (that start with `CC`) were [first published in 2002][cc-v1].
Since then, the possibility to relicense under later ([v2.0][cc-v2], 2004)
and compatible licenses ([v3.0][cc-v3], 2007) has been added,
a 30-day window to correct license violations has been established to combat [copyleft trolls][commons-trolls],
and _sui generis_ database rights have been covered explicitly ([v4.0][cc-v4], 2013).
There are many more [subtle differences between license versions][cc-versions],
therefore it is important to indicate which license version exactly one is referring to,
as the license of a work does not "update" automatically
(i.e., you have to follow the terms of that exact version).
CC0, officially introduced in 2009, is still at version [1.0][cc0-v1].

[cc-v1]: https://creativecommons.org/weblog/entry/3476
[cc-v2]: https://creativecommons.org/weblog/entry/4216#:~:text=Share%20Alike%20Across%20Borders
[cc-v3]: https://creativecommons.org/weblog/entry/7249#:~:text=BY-SA%20%E2%80%94%20Compatibility%20Structure%20Now%20Included
[cc-v4]: https://creativecommons.org/weblog/entry/40768
[cc0-v1]: https://www.creativecommons.org/weblog/entry/13304
[cc-versions]: https://wiki.creativecommons.org/wiki/License_Versions
[commons-trolls]: https://commons.wikimedia.org/wiki/Commons:Copyleft_trolling

For the GNU licenses it is even recommended to state whether a work is licensed
_only_ under the indicated version of the license or, alternatively,
also under newer versions of the license [@Stallman2022Version].
:::

### Under Which Conditions?

Finally, if you verified that the work has been made available to you under a free/open license,
you need to understand the conditions under which you are allowed to include the work in your project folder
and share it with others -- regardless of whether you modified it.
In the following, we will explain the most important conditions of the licenses listed above.
For other licenses we recommend reading their text in full and/or consulting with a lawyer.
The conditions of many popular free/open licenses are also [summarized on ChooseALicense.com][cal-appendix],
which you can use to get a general idea of a license
(but which does not obviate reading the license text in full).
Specifically for software licenses the _Open Source Automation Development Lab_
provides [useful checklists detailing the respective obligations][osadl].[^osadl-note]

[cal-appendix]: https://choosealicense.com/appendix/
[osadl]: https://www.osadl.org/Checklists/
[^osadl-note]: Accessing the checklists requires a (free) account on their website.

::: {#nte-private-use .callout-note}
#### Private Use

In the following, we only discuss the conditions under which works may be _shared_.
This is because free/open licenses put no conditions on private use,
where the work is not made available to others.
:::

#### Attribution

Licenses may require that you provide attribution to the author(s)[^others]
and to retain any copyright notices,[^example-notice]
although the specifics vary between licenses.
For example, Creative Commons licenses other than CC0 require
that the name of the author, the title of the work, and a link to the work are provided.

[^others]: or others whose names are provided with the original work
[^example-notice]: Such as `Copyright (c) 2025 Jane Doe`

Projects with many authors often explicitly allow for an attribution
to the collective as a whole rather than being mentioned individually.
For example, the flexible attribution requirements of Creative Commons licenses
allow that Wikipedia authors agree to be attributed through a link to the article [@Linksvayer2009].
And software projects may use copyright notices that only contain the project name,
such as `Copyright (c) 2020-2025 The XYZ Project Contributors` [@Winslow2020].

Where to put the attribution depends on the type of the work,
but in general it is best to keep it close to where the work is used
and to provide all required information.
Only linking to a page that provides the attribution (as with Wikipedia articles)
most likely will not suffice in general.

#### State Changes

If a work has been modified, licenses may require that this is indicated.
For example, if an image under a Creative Commons license other than CC0
is reused in a cropped version, reusers could add "adapted from original"
to indicate that changes have occurred.

#### Indicate License / Add License Text

Licenses may either require that one mentions the applicable license
by name and possibly link ("Indicate license")
or they even require to copy the full license text along with the used work ("Add license text").
Of course, the latter may not be feasible
if the work is printed and smaller than the license text.
For example, if the license of an image requires to print the full license text,
the image mostly cannot be used in newspapers [see @Wikipedia2024FDL].

#### Disclose Code

There are various situations in which one interacts with software
without having access to the underlying source code.
For example, software can be distributed as a binary ("executable") file
or it can be used over the internet.
In both cases, licenses may require that the source code is provided to its users.

#### Anti-DRM

You may have encountered images or videos which you could not download due to a copy protection.
Such a protection is also called **d**igital **r**ights **m**anagement (DRM).
For works published under a license with an anti-DRM clause,
reusers may not apply DRM to the work or must provide a DRM-free version in parallel
(depending on the exact license).

#### Copyleft

If you modified the work and put a sufficient amount of creativity into it,
the result is called a _derivative work_.
It is (again) protected by copyright and even if you share it,
nobody except you is allowed to copy, modify, or distribute it further.
You can put it under a free/open license but don't have to.[^still-abide]
The only exception is if the original work has been made available under a "copyleft" license.
In that case, if you share the work with others
you must put it under the same license as the original work[^or-compatible]
-- or you commit a copyright violation.[^not-automatic]
This explains the origin of the term _copyleft_,
as it "turns copyright around" to make works permanently free: ©\ →\ 🄯.
In the realm of Creative Commons licenses, copyleft is also called _share alike_.
The opposite of a copyleft license is sometimes also called a _permissive_ license.

[^still-abide]: Of course, you still need to abide
by the license conditions of the original work, such as attribution.

[^or-compatible]: or sometimes also under a later version or under a compatible license

[^not-automatic]: Note, however, that [copyleft licenses are not automatic][commons-automatic]:
A derivative work of a copylefted work is not "implicitly" under the same license
-- others would commit a copyright violation if they just took it.
Only if the derivative author explicitly makes it available under the same license,
others may use the derivative work under the same terms.

[commons-automatic]: https://commons.wikimedia.org/wiki/Commons:Viral_licenses_are_not_automatic

What is considered creation of a derivative work varies between jurisdictions,
but the following are typical examples:
translating a text into a different language,
taking a picture of an artwork,
adding a song to a video, adapting a computer program to fit own use cases,
or taking a [screenshot of a computer program][commons-screenshots].
In all of these cases, if the original work is under a copyleft license,
then sharing of the resulting derivative work is only allowed
if it is put under the same license.

[commons-screenshots]: https://commons.wikimedia.org/wiki/Commons:Screenshots

The question is then: Which uses of a work do not create a derivative work?
In many cases, if an unaltered excerpt "is used to illuminate an idea
or provide an example in another larger work" [@CC2024NoDerivs] no derivative work is made.
The excerpt needs to "remain separate and distinct in the new context" [@Kreutzer2024, p. 44]
and not be "merged with other material into a new and larger work" [@Kreutzer2024, p. 45].
For example, embedding images in a document and clearly separating them from the rest
or assembling works as part of a collection typically would not create a derivative work.
However, due to a lack of case law, there is only little guidance
on what constitutes a derivative work in general.

Also note that the copyleft licenses we discuss here do not mandate sharing.
Copyleft (and attribution) clauses are only triggered if the work is shared [@CC2015].
This means that if you only use a work internally, you do not need to share your derivative works.

::: {#nte-weak-strong-copyleft .callout-note collapse="true"}
##### Weak and Strong Copyleft

For software, copyleft licenses come in two flavors:

- __Weak copyleft licenses__ only require that modifications to the software itself
  are licensed under the same or a compatible license if shared.
  For example, if you create and publish software under a weak copyleft license,
  others who modify it and put their version on the internet have to apply the same license.
  However, people who merely _use_ your software in their own work which they make publicly available
  can choose any license for it.
- __Strong copyleft licenses,__ on the other side, insist that any larger works
  that use the copyleft-licensed software must also be licensed
  under the same or a compatible license if shared.
  For example, if your software were to be put under a strong copyleft license,
  everybody publishing software that uses your software would need to put it under the same license.
  Because of only few rulings in courts, the extent of this requirement is disputed [@Wikipedia2024GPL].
:::

### Further Remarks

The description of the conditions above is a simplified summary
-- read the full text of the license before sharing the respective work with others.
For example, Creative Commons licenses other than CC0 also require that

- you retain a notice about the disclaimer of warranties,
- you retain an indication of any previous modifications,
- the attribution does not imply endorsement by the original author, and
- you remove the attribution again upon request by the original author.

::: {#tip-cc-licenses-checklist .callout-tip}
#### Required Information under Creative Commons Licenses

For most Creative Commons licenses,
providing the following information is required (if available):

- [ ] Author name
- [ ] Copyright notice
- [ ] Title of the work
  - [ ] Link to the work
- [ ] License name
  - [ ] Link to the license
- [ ] Link to warranty disclaimer
- [ ] Indicate modification
:::

Obviously, if you are in doubt whether a condition of the license applies in your use case,
it is better to either follow it or to ask the author for clarification or explicit permission.
This is especially important given the existence of [_copyleft trolls_][commons-trolls]:
People who distribute works under a free/open license
but send out statutory damages claims after minor violations
without providing reusers an opportunity to fix their error.
In addition, in many European countries attribution may be necessary regardless
due to the author's moral rights.
For these reasons, we recommend that you provide attribution
for all works under a free/open license
-- even if the license does not strictly require it (e.g., CC0).
Also, you should take note of every foreign work
included in your project folder and record its license.

::: {#imp-non-copyright-restrictions .callout-important}
#### Non-copyright Restrictions

The free/open licenses we discuss here mostly grant copyrights
(with some of them explicitly granting patent rights and _sui generis_ database rights).
Therefore, you may lack other rights such as personality, privacy, moral, or trademark rights.
For example, [sharing photos that depict people][commons-photographs]
is not only a matter of copyright, but also of privacy rights.
Also, their commercial use may require the consent of the depicted person.
Finally, note that even if attribution is not a requirement of the license,
good scientific practice demands that appropriate citations are made.
:::

[commons-photographs]: https://commons.wikimedia.org/wiki/Commons:Photographs_of_identifiable_people

## Wrap-up

If you would like to learn more about copyright and licenses
you might find the following resources interesting:

- "Open Content -- A Practical Guide to Using Creative Commons Licences"
  by @Kreutzer2024
- "Creative Commons Certificate for Educators, Academic Librarians,
  and Open Culture" by @CC2024Certificate
- "Freie Software -- Zwischen Privat- und Gemeineigentum" by @Grassmuck2004
- "Rechtsfragen bei Open Science: Ein Leitfaden" by @Kreutzer2021