### 2025-07-15 data.qmd
Shorten the data chapter

---
title: "Add Data and Data Dictionary"
engine: knitr
---

## Add Data File

```{r}
#| include: false

invisible(loadNamespace("palmerpenguins")) # Tell renv that we need this package
stopifnot(packageDescription("palmerpenguins", fields = "License") == "CC0")

# write both in the top folder and in the in-depth folder
# (so that both scripts can reference the data without relative paths)
write.csv(palmerpenguins::penguins, file = "data.csv", row.names = FALSE)
write.csv(palmerpenguins::penguins, file = "in_depth_material/data.csv", row.names = FALSE)
```

You can now download the data set we have prepared for you
and put it into your project folder:

[data.csv](data.csv){.btn .btn-lg .btn-info download="data.csv"}

::: {.column-margin}
![palmerpenguins: Palmer Archipelago (Antarctica) Penguin Data](images/palmerpenguins.png){width=250px}
:::

The data set is from the package [`palmerpenguins`][palmerpenguins]
(v`r packageVersion("palmerpenguins")`)
and contains the recorded bill lengths and sex of penguins
living on three islands in the Palmer Archipelago, Antarctica.
It was made available by Allison Horst, Alison Hill, and Kristen Gorman
under the license [CC0\ 1.0][cc0].

[palmerpenguins]: https://allisonhorst.github.io/palmerpenguins/
[cc0]: https://creativecommons.org/publicdomain/zero/1.0/

::: {#imp-legal-restrictions .callout-important}
### Consider Legal Restrictions Before Sharing

Everything you put into the project folder will be shared publicly.
For reasons of reproducibility, this should include the data you analyze.
Of course, you should only share them to the extent you are allowed to.
Besides copyright and similar rights, you need to take into account
applicable privacy laws (e.g., the GDPR for European citizens)
and contractual obligations (e.g., with your data provider).

Privacy laws and contractual obligations may require you
to create an anonymized or synthetic data set[^anon-synth-tools],
in which case you should provide a reference
to a repository where the originally measured data can be obtained from.
For further information, you can watch the talk "[Data anonymity](https://osf.io/j6fy8)"
by Felix Schönbrodt recorded during the LMU Open Science Center Summer School 2023
and have a look at [the accompanying slides](https://osf.io/z6gcu).
Another resource to look at is the presentation
"Creating reproducible packages when data are confidential" by @Vilhuber2024.
:::

[^anon-synth-tools]: For example, using [Amnesia](https://amnesia.openaire.eu/),
[ARX](https://arx.deidentifier.org/), [sdcTools](https://github.com/sdcTools),
[Synthpop](https://www.synthpop.org.uk/), or [OpenDP](https://opendp.org/).

## Add a Data Dictionary

Whether or not distributing the data set,
it is important to document the meaning (e.g., units)
and values of its variables.
This is typically done with a _data dictionary_ (also called a _codebook_).

The recommendation for data dictionaries vary between fields - 
both in terms of the recommended content (i.e., what exactly should be documented) 
and the technical implementation (i.e., which file formats should be used).

*For the purpose of this exercise, we keep it easy and propose that you 
manually create a file with the data dictionary
(e.g., as a table in .xlsx, .docx, .ods, or as a Markdown table), 
documenting only the bare minimum.*

<!-- TODO: Add a link to the forthcoming data dictionary OSC tutorial -->

::: {#tip-in-depth .callout-tip collapse="false"}
### Manual is too much work? Automatic generation of (machine-readable) data dictionaries
Our in-depth supplementary material "[Automatic Generation of Data Dictionaries](in_depth_material/data_dic_generation.qmd)" explains how you can automatically create
a data dictionary with an R package. The package reads the data set and extracts relevant information from it. 
This approach is in particular useful if you need to document data sets with many variables.

This advanced chapter also contains a section on how to create machine-readable data dictionaries.
:::

### A bare minimum data dictionary

Most standards for data dictionaries require at least 
this information for each variable in your data set:

- `name`: The (machine‑readable) name of the variable  
- `label`: A short, human‑readable title or label for the variable  
- `type`: The data type of the variable (e.g., `integer`, `float`, `string`, `date`)  
- `description`: A brief description of what the variable measures or represents  
- `values` (for categorical variables): A mapping of codes to their meanings, for example:  1 = Male, 2 = Female, 9 = Missing)
- `units` (for numeric measures): The units of measurement (e.g., `kg`, `USD`, `years`, or the scale of a survey response), 
- `missing_codes`: Any special codes used to denote missing or non‑applicable values (e.g., `-99 = Not answered`)

Here's an example from a different data set, with variables in rows, and the dictionary in columns:

| name              | label                     | type    | description                             | values                                                                                  | units       | missing_codes      |
| ----------------- | ------------------------- | ------- | --------------------------------------- | --------------------------------------------------------------------------------------- | ----------- | ------------------ |
| gender            | Gender                    | integer | self‑identified gender                  | 1 = Male; 2 = Female; 3 = Other; 9 = Missing                                            |             | 9 = Missing        |
| age               | Age                       | integer | age in years                            |                                                                                         | years       | -99 = Not answered |
| blood_pressure    | Blood Pressure (systolic) | integer | systolic blood pressure                 |                                                                                         | mmHg        | -99 = Not measured |
| life_satisfaction | Life Satisfaction         | integer | "How satisfied are you with your life?" | 1 = Very dissatisfied; 2 = Dissatisfied; 3 = Neutral; 4 = Satisfied; 5 = Very satisfied | scale (1–5) | -99 = Not answered |

### ✍️ Practical Exercise: Add your own data dictionary

Now go ahead and create a data dictionary for the penguins data set, 
in a software (text or spreadsheet) of your choice.

Save the data dictionary in the same folder as the actual data set file.

::: {#tip-data_dic .callout-tip collapse="true"}
#### Data dictionary for penguins data set (Solution)

| name              | label          | type    | description                          | values                    | units | missing_codes |
| ----------------- | -------------- | ------- | ------------------------------------ | ------------------------- | ----- | ------------- |
| species           | Species        | string  | Penguin species                      | Adelie; Gentoo; Chinstrap |       | NA = Missing  |
| island            | Island         | string  | Island where individual was observed | Torgersen; Biscoe; Dream  |       | NA = Missing  |
| bill_length_mm    | Bill length    | float   | Length of the bill (beak)            |                           | mm    | NA = Missing  |
| bill_depth_mm     | Bill depth     | float   | Depth of the bill (beak)             |                           | mm    | NA = Missing  |
| flipper_length_mm | Flipper length | integer | Length of the flipper                |                           | mm    | NA = Missing  |
| body_mass_g       | Body mass      | integer | Body mass                            |                           | g     | NA = Missing  |
| sex               | Sex            | string  | Sex of the penguin                   | male; female              |       | NA = Missing  |
| year              | Year           | integer | Year of data collection              | 2007; 2008; 2009          |       |               |

:::

## ✍️ Practical Exercise: Add Data Citation and Attribution

All data relied upon should be cited in the manuscript
to allow for precise identification and access.
Now, it's your turn to add an appropriate citation for the data set to the manuscript.

Hints:

- You can find an appropriate BibTeX entry on [the package website](https://allisonhorst.github.io/palmerpenguins/#citation)
  or with the function `citation()`^[Note that this function requires to have the respective package installed.].
- Add the citation in the manuscript where it says "cite data here".

::: {#tip-cite-palmerpenguins .callout-tip collapse="true"}
#### Citing the Data Set (Solution)

Show the correct reference of the data set:

```{r}
#| class.output: "bib code-overflow-scroll"
#| attr-output: 'filename="Bibliography.bib"'

citation("palmerpenguins", auto = TRUE) |>
  transform(key = "horst2020") |>
  toBibtex()
```

Copy the BibTeX entry to the file `Bibliography.bib`.
Then, find the line in the manuscript that says "cite data here"
and replace it with a sentence such as the following:

```{.md filename="Manuscript.qmd"}
The analyzed data are by @horst2020.
```

Render the document to check that the citation is displayed properly.

```{.bash filename="Terminal"}
quarto render Manuscript.qmd
```
:::

## Wrap up

Congrats! You documented your data set and cited it correctly.

To finalize this step, you can go through the commit routine:

```{.bash filename="Terminal"}
git status
git add .
git commit -m "Add data"
```